### 2017-02-28 First_np_Paper/first_np_paper.tex
Updated section 2.7 with Lee edits

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Important note on usage                  %%
%% -----------------------                  %%
%% This file must be compiled with PDFLaTeX %%
%% Using standard LaTeX will not work!      %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The '3p' and 'times' class options of elsarticle are used for Elsevier CRC
\documentclass[5p]{elsarticle}
% \documentclass[3p,times]{elsarticle}
% \documentclass[3p]{elsarticle}

\usepackage[american]{babel}
\usepackage{amsmath}
\usepackage[version=3]{mhchem} 
% \usepackage{fixltx2e}
% \usepackage{refcount}
% \usepackage{siunitx}
% \usepackage{lastpage}
% \usepackage{textcomp}
\usepackage{mathtools}

\usepackage{xfrac}
\usepackage{lmodern}
\usepackage[hidelinks]{hyperref}
% \usepackage{cool}
% \usepackage{cancel}
\usepackage{microtype}
\usepackage{listings}
\usepackage{mcode}
\usepackage [autostyle, english = american]{csquotes}
\usepackage{longtable}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{gensymb}

% \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{color}

\usepackage[colorinlistoftodos]{todonotes}

\usepackage[section]{placeins}

\lstset{basicstyle=\small\ttfamily,columns=fullflexible}

% \usepackage{verbatim}

% \usepackage{gensymb}
% \usepackage{enumerate}
% \usepackage{float}
% \usepackage{bm}
% \usepackage{csquotes}
% \usepackage{mathtools}
% \usepackage{natbib}
% \usepackage{biblatex}

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}

%% The ecrc package defines commands needed for running heads and logos.
%% For running heads, you can set the journal name, the volume, the starting page and the authors

%% set the volume if you know. Otherwise `00'
\volume{00}

%% set the starting page if not 1
\firstpage{1}

%% Give the name of the journal
\journalname{Nuclear Instruments and Methods in Physics Research B}

%% Give the author list to appear in the running head
%% Example \runauth{C.V. Radhakrishnan et al.}
\runauth{A.S. Voyles et al.}

%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo

%% Give the abbreviation of the Journal.
\jid{nimb}
% \jid{yspmi}

%% Give a short journal name for the dummy logo (if needed)
\jnltitlelogo{Nucl Instrum Meth B}

%% Hereafter the template follows `elsarticle'.
%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.

%% Elsevier CRC generally uses a numbered reference style
%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
%% If using BibTeX, use the style file elsarticle-num.bst

%% End of ecrc-specific commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
\usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

% if you have landscape tables
\usepackage[figuresright]{rotating}

% put your own definitions here:
%   \newcommand{\cZ}{\cal{Z}}
%   \newtheorem{def}{Definition}[section]
%   ...

% add words to TeX's hyphenation exception list
%\hyphenation{author another created financial paper re-commend-ed Post-Script}

% declarations for front matter

\usepackage{fancyvrb}
\usepackage{color}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{ %
  backgroundcolor=\color{white},   % choose the background color
  basicstyle=\footnotesize,        % size of fonts used for the code
  breaklines=true,                 % automatic line breaking only at whitespace
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  keywordstyle=\color{blue},       % keyword style
  stringstyle=\color{mymauve},     % string literal style
}

% Sin and Cos with auto-parentheses 
\newcommand{\sinp}[1]{\sin{\left( #1\right)}}
\newcommand{\cosp}[1]{\cos{\left( #1\right)}}
\newcommand{\expp}[1]{\exp{\left( #1\right)}}
\newcommand{\sinhp}[1]{\sinh{\left( #1\right)}}
\newcommand{\lnp}[1]{\ln{\left( #1\right)}}
\newcommand{\pp}[1]{\left( #1\right)}
\newcommand{\sci}[2]{ #1 \cdot 10^{#2}\ }
\newcommand{\angstrom}{\mbox{\normalfont\AA}}
\newcommand{\norm}[1]{\lVert #1 \rVert}

\newcommand{\colornote}[1]{\textcolor{red}{ COMMENT\large\footnote{\textcolor{red}{#1}}}}

\newcommand{\comment}[1]{\todo[color=blue!20!white,inline]{ASV: #1}}

\makeatletter
% Make common definition of mean
\newcommand*\mean[1]{\overline{#1\raisebox{3mm}{}}}

\makeatother

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
\title{Determination of the \ce{^{64}Zn},\ce{^{47}Ti}(n,p) Cross-Sections using a DD Neutron Generator for Medical Isotope Studies}

% \title{(n,p) Cross-Section Measurements for \ce{^{64}Cu} and \ce{^{47}Sc} via DD Neutron Generator\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

% \dochead{Short}
%% Use \dochead if there is an article header, e.g. \dochead{Short communication}

% \title{NE290A - Homework \#04}
% \date{29 April, 2016}

% \author[rvt]{C.V. ̃Radhakrishnan\corref{cor1}\fnref{fn1}}
\author[ucb]{A.S. Voyles \corref{cor1}}
\ead{andrew.voyles@gmail.com}

\author[lbl]{M.S. Basunia}

\author[lbl]{J.C. Batchelder}

\author[llnl]{J.D. Bauer}

\author[ucb,lbl]{L.A. Bernstein}

\author[ucb]{E.F. Matthews}

\author[geo,eps]{D. Rutte}

\author[ucb]{M.A. Unzueta}

\author[ucb]{K.A. van Bibber}

% \author[lbl]{K. ̃Bazargan\fnref{fn2}}
% \ead{kaveh@river-valley.com}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\cortext[cor1]{Corresponding author}
% \cortext[cor2]{Principal corresponding author}
% \fntext[fn1]{This is the specimen author footnote.}
% \fntext[fn2]{Another author footnote, but a little more longer.}

% \address[ucb]{Department of Nuclear Engineering, University of California, Berkeley, Etcheverry Hall, 2521 Hearst Ave, Berkeley, CA 94709}
% \address[lbl]{Lawrence Berkeley National Laboratory,  1 Cyclotron Rd, Berkeley, CA 94720}
% \address[llnl]{Lawrence Livermore National Laboratory, 7000 East Ave, Livermore, CA 94550}

\address[ucb]{Department of Nuclear Engineering, University of California, Berkeley, Berkeley CA, 94720 USA}
\address[lbl]{Lawrence Berkeley National Laboratory,  Berkeley CA, 94720 USA}
\address[llnl]{Lawrence Livermore National Laboratory, Livermore CA, 94551 USA}
\address[geo]{Berkeley Geochronology Center, Berkeley CA,  94709  USA}
\address[eps]{Department of Earth and Planetary Sciences, University of California, Berkeley, Berkeley CA,  94720  USA}

% \author{Andrew Voyles}

% \address{Nuclear Engineering, University of California Berkeley, Berkeley, CA 94709, USA.}

\begin{abstract}

%\comment{Use DD or DD as convention?}

% \comment{Need Daniel's middle initial} - he doesn't have one

Cross sections for the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} and \ce{^{64}Zn}(n,p)\ce{^{64}Cu} reactions have been measured for quasi-monoenergetic DD neutrons produced by the UC-Berkeley High Flux Neutron Generator. The study was motivated by interest in the production of \ce{^{47}Sc} and \ce{^{64}Cu} as emerging medical isotopes. The cross sections were measured in ratio to the \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} inelastic scattering reactions on co-irradiated indium samples. Post-irradiation counting using an HPGe and LEPS detectors allowed for cross section determination to within 6\% uncertainty. The cross section were determined with lower uncertainty than existing measurements and are found to be  in good agreement with the literature and theoretical values. This work highlights the utility of using DD plasma-based neutron sources, for a host of nuclear data measurements and potentially for the production of radionuclides for medical applications.

\comment{The abstract is now nicely short and sweet, but should it be expanded at all?}

\comment{For my corresponding author email, should I list my Gmail (more permanent) or my Berkeley / LBL address (more professional)?}

\end{abstract}

% \begin{keyword}
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

% \end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers

\listoftodos

%% main text 
\section{Introduction} \label{sec:intro}

%
% If we want to mention in-situ-production
%
%
% e.g.:In-situ production (I always figured this is what is so great, correct me if I am wrong) of radioactive nuclides is an emerging method for radiation therapy in nuclear medicine (e.g., cite some recent review paper). It is based on doping the malignant target tissue with a stable nuclide which is then activated in-situ by neutron irradiation. Following decay is focused on the target tissue. Advantages compared to conventional radiation therapy are the reduction of damage to surrounding cells, … short range, … . Currently development of this technique is hampered by the lack of fundamental cross section data for various isotopes.

There has been significant interest in the past several years in exploring the use of neutron-induced reactions to create radionuclides for a wide range of applications, due to the far greater range of neutrons as compared to charged particle beams ($\sfrac{\text{g}}{\text{cm}^2}$ as compared to 10's of $\sfrac{\text{mg}}{\text{cm}^2}$), together with the fact that isotope production facilities often produce large secondary neutron fields.  Particular interest has been paid to (n,p) and (n,$\alpha$) charge-exchange reactions since these reactions produce high-specific activity radionuclide samples without the use of chemical carriers in the separation process.

Two other potential neutron sources for (n,x) reactions exist in addition to the secondary neutron fields generated at existing isotope production facilities: reactors and neutron generators that utilize the D(T,n)$\alpha$ \enquote{DT} and D(D,n)\ce{^3He} \enquote{DD} reactions.  While reactors produce copious quantities of neutrons, their energy spectrum is often not well-suited to the preparation of high-purity samples due to the co-production of unwanted activities via neutron capture, in addition to the significant start-up costs and proliferation concerns involved in their commissioning.  Similarly, while the higher energy 14-15 MeV neutrons produced at DT generators are capable of initiating (n,p) and (n,$\alpha$) reactions, their higher energy opens the possibility of creating unwanted activities via (n,pxn) and (n,$\alpha$xn) reactions that cannot easily be separated from the desired radionuclides.

In contrast, while the neutron spectrum from a DD reaction, which ranges from approximately 2-3 MeV, is ideally suited to radionuclide production, the lower flux from these generators limits their production rate.  An additional complication is the relative paucity of high-quality, consistent cross section data for neutrons in the 2-3 MeV DD energy range.

The purpose of the present work is to explore the potential for using high-flux neutron generators to make high-specific activity samples of radionuclides at the mCi level for local use in the application community.  The research group at UC-Berkeley has been developing a High Flux Neutron Generator (HFNG) that features an internal target where samples can be placed several mm from the neutron producing surface in order to maximize the utilization of each neutron for the production of a desired radionuclide \textcolor{red}{[XXXref Cory paper]}.  The HFNG uses the D(D,n)\ce{^3He} reaction to produce neutrons with energies near 2.45 MeV together with a self-loading target design to maintain continuous operation without target replacement.  In addition to the generator itself, efforts are underway to design neutron reflection capabilities to allow scattered neutrons multiple opportunities to interact with an  internally mounted target. While these design efforts are underway, the HFNG can be used to better characterize production cross sections at the appropriate neutron energy.

\comment{cite Cory's paper}

The present work features a pair of cross section measurements for the production of two non-standard positron emitters: \ce{^{64}Zn}(n,p)\ce{^{64}Cu} and \ce{^{47}Ti}(n,p)\ce{^{47}Sc}. \ce{^{64}Cu}  ($t_{1/2}$ = 12.7 h) undergoes $\beta^+$ decay (61.5\% branching ratio) to \ce{^{64}Ni} or $\beta^-$ decay (38.5\% branching ratio) to \ce{^{64}Zn}. The emitted 190-keV $\beta^-$ particle makes this an  attractive short-range therapeutic radionuclide, which also has the possibility for simultaneous positron emission tomography (PET) imaging for real-time dose monitoring and verification. This makes \ce{^{64}Cu} particularly attractive for emerging radiation therapy protocols. In addition, copper radiochemistry is well developed, and many existing ligands and carriers may be used for selective delivery of the radionuclide to different sites in patients. The second radionuclide studied, \ce{^{47}Sc} ($t_{1/2}$ = 3.35 d), undergoes $\beta^-$ decay to \ce{^{47}Ti}, emitting a high-intensity 159-keV gamma ray in the process. This radionuclide is  attractive as an emerging diagnostic isotope, due to the similarity of the emitted gamma ray to that of the more well-established \ce{^{99m}Tc}. However, due to the short half-life of and dwindling supplies of \ce{^{99m}Tc}, \ce{^{47}Sc} stands poised as a potential solution to this shortage, due to its longer half-life and multiple production pathways without the need for highly enriched uranium. In addition, when paired with \ce{^{44}Sc}, \ce{^{47}Sc} forms a promising \enquote{theranostic} set for use in simultaneous therapeutic and diagnostic applications.

\comment{cite theranostic papers, etc}

Current methodology in radiochemistry has been proven to successfully recover upwards of 95\% of produced \ce{^{64}Cu} and \ce{^{47}Sc} from solid target designs, without the need for additional carrier. By expanding the base of efficient reaction pathways, great advances can be made in making production of medical radionuclides more efficient and affordable for those in need.

\section{Experiment}\label{sec:experiment}

\subsection{Neutron source}\label{sec:n_source}

\comment{Regex to replace all hard figure references with LaTeX cross-references}

% Neutron activation was carried out via irradiation in the High-Flux Neutron Generator (HFNG), a DD neutron generator at the University of California, Berkeley, seen in  \autoref{fig:hfng_a}. This generator extracts deuterium ions from an RF-heated deuterium plasma through a nozzle, whose shape was designed to form a flat-profile beam 5 mm in diameter. This deuterium beam is incident upon a water-cooled, self-loading titanium-coated copper target, where the titanium layer acts as a reaction surface for DD fusion, producing neutrons with a well-known energy distribution as a function of  emission angle \cite{Liskien_Paulsen_1973}. Targets are inserted in the center of the target, approximately 8 mm from the DD reaction surface, prior to startup. \autoref{fig:hfng_b} displays a cut-away schematic of the HFNG. A 100 kV deuterium beam was extracted at 1.5 mA, creating a fluence of approximately  $\sci{1.3}{7}$ neutrons/cm$^2$ on the target.

Neutron activation was carried out via irradiation in the High-Flux Neutron Generator (HFNG), a DD neutron generator at the University of California, Berkeley. This generator extracts deuterium ions from an RF-heated deuterium plasma through a nozzle, whose shape was designed to form a flat-profile beam 5 mm in diameter. This deuterium beam is incident upon a water-cooled, self-loading titanium-coated copper target, where the titanium layer acts as a reaction surface for DD fusion, producing neutrons with a well-known energy distribution as a function of  emission angle \cite{Liskien_Paulsen_1973}. Targets are inserted in the center of the target, approximately 8 mm from the DD reaction surface, prior to startup. \autoref{fig:hfng_b} displays a cut-away schematic of the HFNG. A 100 kV deuterium beam was extracted at 1.3 mA, creating a flux of approximately  $\sci{1.3}{7}$ neutrons/cm$^2$s on the target.

% \begin{figure}
%  \centering
% %  \includegraphics[width=\columnwidth]{./figures/IMG_20160531_183154957_HDR.jpg}
%  \includegraphics[scale=0.06]{./figures/IMG_20160531_183154957_HDR.jpg}
%  % IMG_20160531_183154957_HDR.jpg: 2432x4320 pixel, 72dpi, 85.80x152.40 cm, bb=0 0 2432 4320
%  \caption{Alternate HFNG picture - \color{red}{use this or  \autoref{fig:hfng_a}?}}
%  \label{fig:alt_HFNG}
% \end{figure}

% \begin{figure*}
%     \centering
%     \begin{subfigure}[t]{0.4\textwidth}
%         \centering
% %         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
%         \includegraphics[height=3in]{./figures/Capture.PNG}
%         \caption{The UC-Berkeley High-Flux Neutron Generator,}
%         \label{fig:hfng_a}
%     \end{subfigure}%
%     ~ 
%     \begin{subfigure}[t]{0.4\textwidth}
%         \centering
% %         \includegraphics[width=\textwidth]{./figures/target2.png}
%         \includegraphics[height=2in]{./figures/target2.png}
%         \caption{Cutaway of the target chamber.}
%                 \label{fig:hfng_b}
%     \end{subfigure}
%     \caption{Note: the ion source is approximately 20 cm in diameter.}
%      \label{fig:main_HFNG}
% \end{figure*}

\begin{figure}
    \centering
        \centering
%         \includegraphics[width=\textwidth]{./figures/target2.png}
        \includegraphics[height=2in]{./figures/target2.png}
        \caption{Cut-away schematic of the HFNG. Note: the ion source is approximately 20 cm in diameter.}
                \label{fig:hfng_b}
%      \label{fig:main_HFNG}
\end{figure}

\subsection{Cross section determination by relative activation}\label{sec:sample_loading}

The approach used in both measurements was to irradiate foils of zinc or titanium, which were co-loaded with indium foils in order to determine their (n,p) cross sections relative to the well-established \ce{^{113}In}(n,n')\ce{^{113m}In} and \ce{^{115}In}(n,n')\ce{^{115m}In} neutron dosimetry standards.   \autoref{tab:foil_specs} records physical characteristics of each foil for the various irradiations. In each experiment, the co-loaded foils were irradiated for 3 hours at nominal operating conditions of 1.3 mA and 100 kV.  After irradiation, the foils were removed and placed in front of an appropriate High-Purity Germanium gamma-ray detector and the decay gamma-rays were recorded in variable-time bins.

\begin{figure}
    \centering
%         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
        \includegraphics[height=2in]{./figures/holder.pdf}
        \caption{Schematic (not drawn to scale) of the sample holder used for the Berkeley HFNG,}
        \label{fig:holder_a}
\end{figure}

% \begin{figure*}
%     \centering
%     \begin{subfigure}[t]{0.3\textwidth}
%         \centering
% %         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
%         \includegraphics[height=2in]{./figures/foils.png}
%         \caption{Schematic (not drawn to scale) of the foil co-loading holder for the Berkeley HFNG,}
%         \label{fig:holder_a}
%     \end{subfigure}%
%     ~ 
%     \begin{subfigure}[t]{0.3\textwidth}
%         \centering
% %         \includegraphics[width=\textwidth]{./figures/target2.png}
%         \includegraphics[height=2in]{./figures/holder.pdf}
%         \caption{\color{red}{Use this or \autoref{fig:holder_a} - which is more appropriate?}}
%                 \label{fig:holder_b}
%     \end{subfigure}
%      ~ 
%     \begin{subfigure}[t]{0.3\textwidth}
%         \centering
% %         \includegraphics[width=\textwidth]{./figures/target2.png}
%         \includegraphics[height=2in]{./figures/IMG_20151103_113432563.jpg}
%         \caption{ zinc (visible) foil co-loaded on top of reference indium foil, ready for irradiation.}
%                 \label{fig:holder_c}
%     \end{subfigure}
%     \caption{}
%      \label{fig:main_holders}
% \end{figure*}

1 cm diameter, 1 mm thick natural abundance zinc and titanium targets were employed for the measurement. Each of these was  co-loaded with a natural abundance Indium foil of 1 cm diameter and 0.5 mm thickness in a recess cut into a 2-mm thick polyethylene holder, as seen in \autoref{fig:holder_a}, which was mounted in the HFNG target center. Prior to loading, each foil was washed with isopropanol and dried, to remove any trace oils or residue that could become activated during irradiation.

\comment{Regex to replace table hard link with LaTeX cross-reference.}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
% \usepackage{graphicx}
\begin{table*}
\centering
\caption{Foil characteristics for each of the three (Zn/In)* experiments and the two (Ti/In)$^\dagger$  experiments. 
}
\label{tab:foil_specs}
\resizebox{\textwidth}{!}{%
\begin{tabular}{@{}ccccccc@{}}
\toprule
Foils Used & Metal Purity          & Abundance (a/o)                                                           & \begin{tabular}[c]{@{}c@{}}Foil Density \\ (mg/cm$^2$)\end{tabular} & Thickness (mm)                                                                                                         & Diameter (cm)                                                                                                           & Mass (g)                                                                                    \\ \midrule
$^\text{nat}$In      & \textgreater 99.999\% & \begin{tabular}[c]{@{}c@{}}\ce{^{113}In} (4.29\%),\\ \ce{^{115}In} (95.71\%)\end{tabular} & 365.5                                                            & \begin{tabular}[c]{@{}c@{}}0.49 $\pm$ 0.02*, \\ 0.50 $\pm$ 0.03*, \\ 0.49 $\pm$ 0.03*, \\ 0.53 $\pm$ 0.06$^\dagger$, \\ XXX $\pm$ XXX$^\dagger$\end{tabular} & \begin{tabular}[c]{@{}c@{}}9.75 $\pm$ 0.09*, \\ 9.98 $\pm$ 0.15*, \\ 9.96 $\pm$ 0.10*, \\ 10.01 $\pm$ 0.11$^\dagger$, \\ XXX $\pm$ XXX$^\dagger$\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.248*, \\ 0.248*, \\ 0.241*, \\ 0.247$^\dagger$,\\  XXXXXX$^\dagger$\end{tabular} \\
$^\text{nat}$Zn      & \textgreater 99.99\%  & \ce{^{64}Zn} (49.17\%)                                                            & 714.1                                                            & \begin{tabular}[c]{@{}c@{}}1.03 $\pm$ 0.01, \\ 1.03 $\pm$ 0.01, \\ 1.02 $\pm$ 0.01\end{tabular}                                    & \begin{tabular}[c]{@{}c@{}}9.93 $\pm$ 0.14, \\ 9.76 $\pm$ 0.17, \\ 9.89 $\pm$ 0.15\end{tabular}                                     & \begin{tabular}[c]{@{}c@{}}0.538, \\ 0.451, \\ 0.452\end{tabular}                           \\
$^\text{nat}$Ti      & 99.999\%              & \ce{^{47}Ti} (7.44\%)                                                             & 450.6                                                            & \begin{tabular}[c]{@{}c@{}}1.16 $\pm$ 0.02, \\ XXXXXX\end{tabular}                                                         & \begin{tabular}[c]{@{}c@{}}9.93 $\pm$ 0.04, \\ XXXXXX\end{tabular}                                                          & \begin{tabular}[c]{@{}c@{}}0.337, \\ XXXXXXX\end{tabular}                                   \\ \bottomrule
\end{tabular}%
}
\end{table*}

\comment{Need foil data from Mauricio's 2nd Ti run}

\subsection{Determination of effective neutron energy}\label{sec:neutron_energies}

The D(D,n)\ce{^3He} reaction at 100 keV lab energy produces neutrons with energies ranging from 2.2 to 2.85 MeV, over an angular range of 0-180\degree\ in the lab frame-of-reference with respect to the incident deuteron beam. This distribution has been well documented \cite{Liskien_Paulsen_1973} and is shown in \autoref{fig:scatt_angle} for 100 keV incident deuteron energy.

\comment{These 2.2 to 2.85 MeV numbers do not quite correspond to the Figure. Addressing this with Joe.}

\comment{Is the one good reference (Liskien) sufficient?}

\comment{Should we say more about the neutron angular and energy distributions? Something from Cory's thesis or the NIM-A papers that are coming along?}

% \comment{Parenthetical citations, or just a list at the end? The Shimizu paper uses the latter as citation format.}

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/scatt_angle.pdf}
 % scatt_angle.pdf: 369x295 pixel, 72dpi, 13.02x10.41 cm, bb=0 0 369 295
 \caption{Energy-angle distribution for neutrons emitted following DD fusion, for 100 keV incident deuterons.}
 \label{fig:scatt_angle}
\end{figure}

Since the samples are separated by only  8 mm from the DD reaction surface they  subtend a fairly significant  (~28\degree) angular range and experience an impressive  (approximately $\sci{1.3}{7}$neutrons/cm$^2$ s) neutron flux. This stands in contrast to other measurements which feature collimated beams and significantly lower total neutron flux.

\begin{figure}
 \centering
 %trim option's parameter order: left bottom right top
 \includegraphics[trim = 0mm 0mm 2mm 0mm, clip,width=\columnwidth]{./figures/mcnp_vised2.png}
 % mcnp_vised2.PNG.png: 688x443 pixel, 96dpi, 18.21x11.72 cm, bb=0 0 516 332
 \caption{MCNP6 model of the HFNG target chamber, with reference scale. The co-loaded foils can be seen in the target chamber center.  The ovals indicate the location of water cooling channels.}
 \label{fig:mcnp_vised}
\end{figure}

\comment{Need to replace this figure w/ non-capitalized labels - this figure needs significant work, and I would appreciate anyone who can provide a better one}

The Monte Carlo N-Particle transport code  MCNP6 \cite{goorley2013initial} was used to model the neutron energy spectrum incident upon target foils co-loaded into the HFNG (see \autoref{fig:mcnp_flux}). This spectrum illustrates the forward-focused kinematics of the DD reaction subtended by the co-loaded sample foils.

% \begin{figure}
%  \centering
%  \includegraphics[scale=0.6]{./figures/mcnp_flux.pdf}
%  % mcnp_flux.pdf: 373x295 pixel, 72dpi, 13.16x10.41 cm, bb=0 0 373 295
%  \caption{MCNP6-modeled neutron energy spectrum for the Berkeley HFNG, using co-loaded indium and zinc (for \ce{^{64}Cu} production) foils as an example.}
%  \label{fig:mcnp_flux}
% \end{figure}

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/mcnp_flux_new.pdf}
 % mcnp_flux.pdf: 373x295 pixel, 72dpi, 13.16x10.41 cm, bb=0 0 373 295
 \caption{MCNP6-modeled neutron energy spectrum for the HFNG.  The solid lines show the spectrum at the location of the indium and the target foil respectively. The dotted and dashed lines show the same with the target itself \enquote{voided} to remove scattering contributions.  Less than 1\%  of the neutrons incident on the foils can be attributed to scatter in the neutron production target.}
 \label{fig:mcnp_flux}
\end{figure}

%\comment{Use this version, old version (mcnp-flux.pdf), or some new combination of lines?}

While this shows that the sample foils experience a very narrow energy distribution of incident neutrons, an effective neutron energy window must be determined. The MCNP6 simulation shows an identical flux-weighted average neutron energy of 2.765 MeV for both the Indium and target foils to the 1 keV level. Due to the kinematics of DD neutron emission, $E_{max}$,  the maximum energy of a neutron subtending the target foils in this geometry is 2.783 MeV. For this maximum energy, the number of reactions induced in a foil (containing $N_i$ target nuclei) is given by:

\begin{equation}
R = N_i \int_0^{E_{max}} \sigma(E) \dfrac{d\phi}{dE} dE
\end{equation}

\comment{Regex to replace equations with LaTeX formats, add cross-references.}

From this definition, it is possible to calculate $F\pp{E'}$, the fraction of total reactions induced up to some energy $E' < E_{max}$:

\begin{equation}\label{eqn:react_fraction}
F\pp{E'} = \dfrac{\int_0^{E'} \sigma(E) \dfrac{d\phi}{dE} dE}{\int_0^{E_{max}} \sigma(E) \dfrac{d\phi}{dE} dE}
\end{equation}

% This quantity $F\pp{E'}$ is plotted in \autoref{fig:frac_plot}.      - note: moved to next paragraph, for better flow.

% \begin{figure}
%  \centering
%  \includegraphics[scale=0.6]{./figures/fracplot.pdf}
%  % fracplot.pdf: 372x295 pixel, 72dpi, 13.12x10.41 cm, bb=0 0 372 295
%  \caption{Fraction of total reactions induced in the Indium foil between the energies [0, $E'$]}
%  \label{fig:frac_plot}
% \end{figure}

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/fracplot_new.pdf}
 % fracplot.pdf: 372x295 pixel, 72dpi, 13.12x10.41 cm, bb=0 0 372 295
 \caption{Fraction of total reactions induced in the Indium foil between the energies [0, $E'$]. The red arrows indicate the energy region that corresponds to 67\% of the total activation.}
 \label{fig:frac_plot}
\end{figure}

%\comment{Use this version, or old version (fracplot.pdf)?}

\comment{Change \autoref{fig:frac_plot} text from the $\pm 1\sigma$ to 68.2\%}

This quantity $F\pp{E'}$ is plotted in \autoref{fig:frac_plot}. The fraction of total reactions in the indium foil can be used to characterize the effective neutron energy bin.  Our approach, in analogy to the Gaussian quantity $\sigma$, will be to use a horizontal \enquote{error bar}  to represent the energy range responsible for 68.2\% of the reactions taking place.  Using this approach, we report the effective energy bin as being $E_n$ = 2.7645 MeV + 0.0151 MeV / - 0.0219 MeV.  This 37-keV full-energy spread verifies that, at such close distances to the DD reaction surface, loaded target foils receive a quasi-monoenergetic neutron flux.

% \comment{Truncate sig figs later on, after discussing systematic errors.}

\subsection{Measurement of induced activities}\label{sec:spectroscopy}

After irradiation, the co-loaded targets were removed from the HFNG and transferred to a counting lab, where their induced activities could be measured via gamma ray spectroscopy. Two detectors were used in this measurement. An Ortec 80\% High-Purity Germanium (HPGe) detector was used for the detection of the positron annihilation radiation from the \ce{^{64}Cu}  decay, the 336 keV gamma-ray from the decay of the \ce{^{115m}In}  isomer and the 391 keV gamma-ray from the \ce{^{113m}In}  isomer. An Ortec planar Low-Energy Photon Spectrometer (LEPS)  was used for the detection of the lower-energy 159 keV gamma-ray from \ce{^{47}Sc} as well as the two indium isomers mentioned above. Both detectors were calibrated for energy and efficiency, using \ce{^{133}Ba}, \ce{^{137}Cs}, and \ce{^{152}Eu} sources at various distances from the front face of each detector. These efficiencies, along with gamma ray intensities for each transition, are used to convert the integrated counts in each gamma ray photopeak into an activity for the activated isotopes and isomeric states.

\comment{You should add a table listing the gamma-ray properties for the indium isomers and the 64Cu and 47Sc decay lines. This should include the proper NDS citation.}

% \begin{figure*}[H]
%     \centering
%     \begin{subfigure}[t]{0.7\textwidth}
%         \centering
% %         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
%         \includegraphics[height=2.8in]{./figures/IMG_20160531_182618557_HDR.jpg}
%         \caption{ Ortec 80\% HPGe detector}
%         \label{fig:hpge_a}
%     \end{subfigure}%
%     ~ 
%     \begin{subfigure}[t]{0.28\textwidth}
%         \centering
% %         \includegraphics[width=\textwidth]{./figures/target2.png}
%         \includegraphics[height=2.8in]{./figures/IMG_20160531_182635712_HDR.jpg}
%         \caption{Ortec planar LEPS detector}
%                 \label{fig:hpge_b}
%     \end{subfigure}
%     \caption{High-Purity Germanium Detectors used for gamma spectroscopy of the activated foils.}
%      \label{fig:main_ge_detectors}
% \end{figure*}
% 
% \comment{Does \autoref{fig:main_ge_detectors} serve any purpose? Should it be cut?}

The foils were counted in their polyethylene holder, 10 cm from the front face of the 80\% HPGe and 1 cm from the front face of the LEPS, with the target foil (zinc or titanium) facing towards the front face of the detector when both target and monitor foils were counted simultaneously. All data collection was performed using the Ortec MAESTRO software. For each experiment the detector dead time was verified to be less than 5\%.  No summing corrections need to be made since all of the gammas are either non-coincident or formed in a back-to-back annihilation event. This was never an issue for any experiments, but had it been, the foils would be moved to a position of greater standoff from the detector.

For the  \ce{^{47}Sc} production experiments, the foils were counted simultaneously using a planar LEPS detector. For the \ce{^{64}Cu} production experiments, the Indium foil was first counted separately using an 80\% HPGe detector, to capture the short-lived Indium activities. This is due to the fact that the \ce{^{115}In}(n,$\gamma$) reaction results in the production of \ce{^{116}In} which has a 54 minute half-life and results in the production of 1097 keV (58.5\% branching), 1293 keV (84.8\% branching) and 2112 keV (15.09\% branching) gamma-rays that in turn produce a significant number of 511 keV gammas from pair-production followed by annihilation. The foils were counted together again after approximately 4 hours of separate collection, to allow for nearly all of the \ce{^{116}In} to decay. Example spectra for each production pathway can be seen in \autoref{fig:spectra_a} and \autoref{fig:spectra_b}.

\begin{figure*}[ht]
    \centering
    \begin{subfigure}[t]{\textwidth}
        \centering
%         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
%         \includegraphics[width=5in]{./figures/ti_foils.png}
        \includegraphics[width=5in]{./figures/47sc_gspectrum_alt.pdf}
        \caption{ Gamma spectrum for the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} production pathway foils, counted using a LEPS detector}
        \label{fig:spectra_a}
    \end{subfigure}%
    \\
    \begin{subfigure}[t]{\textwidth}
        \centering
%         \includegraphics[width=\textwidth]{./figures/target2.png}
%         \includegraphics[width=5in]{./figures/Sample_spectra.png}
        \includegraphics[width=5in]{./figures/64Cu_gspectrum_alt.pdf}
        \caption{Gamma spectrum for the \ce{^{64}Zn}(n,p)\ce{^{64}Cu} production pathway foils, counted using an 80\% HPGe detector.}
                \label{fig:spectra_b}
    \end{subfigure}
    \caption{Example gamma spectra collected to monitor radioisotope production.}
     \label{fig:main_spectra}
\end{figure*}

To verify that each peak corresponds to the assigned decay product, spectra were acquired in a sequence of 15 - 30 minute intervals. The resulting time series displayed in Figures \ref{fig:decay_curve_336} - \ref{fig:decay_curve_511} allow the fitting of exponential decay functions for each nuclide and comparison of the measured half-life with literature values. The fitted functions for each transition agree (at the 1$\sigma$ confidence level) with accepted half-lives \textcolor{red}{[XXXref reference original papers for the T1/2s]}, confirming the respective peak assignments.

\comment{Reference original ENSDF papers}

\begin{figure*}
    \centering
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
%         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
        \includegraphics[scale=0.6]{./figures/336keV_curve_new.pdf}
        \caption{ Decay curve for the isomeric transition of \ce{^{115m}In}.}
        \label{fig:decay_curve_336}
    \end{subfigure}%
     \begin{subfigure}[t]{0.49\textwidth}
        \centering
%         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
%         \includegraphics[scale=0.6]{./figures/391keV_curve2.png}
        \includegraphics[scale=0.6]{./figures/391keV_curve_new.pdf}
        \caption{ Decay curve for the isomeric transition of \ce{^{113m}In}.}
        \label{fig:decay_curve_391}
    \end{subfigure}%
    \\
    \begin{subfigure}[t]{0.49\textwidth}
        \centering
%         \includegraphics[width=\textwidth]{./figures/target2.png}
        \includegraphics[scale=0.6]{./figures/417keV_curve2.png}
        \caption{Decay curve for the $\beta^-$ decay of \ce{^{116}In}.}
                \label{fig:decay_curve_417}
    \end{subfigure}
     \begin{subfigure}[t]{0.49\textwidth}
        \centering
%         \includegraphics[width=\columnwidth]{./figures/Capture.PNG}
        \includegraphics[scale=0.6]{./figures/511keV_curve2.png}
        \caption{ Decay curve for the $\beta^+$ decay of \ce{^{64}Cu}.}
        \label{fig:decay_curve_511}
    \end{subfigure}%
    \caption{Decay curves used to verify photopeak transition assignment.}
     \label{fig:decay_curves}
\end{figure*}

The spectra for the different samples were summed and the peak area over background was determined using gf3, part of the RadWare analysis package from Oak Ridge   \cite{radford2000radware, Radford1995}. The background-subtracted integrated counts in each photopeak, as well as the counting duration for each experiment, are tabulated in \autoref{tab:peak_counts}.

% Please add the following required packages to your document preamble:
% \usepackage{graphicx}
\begin{table*}
\centering
\caption{Counting times and photopeak counts for each of the (Zn/In) and (Ti/In)  experiments. \color{red}{Need Mauricio data!}}
\label{tab:peak_counts}
\resizebox{\textwidth}{!}{%
\begin{tabular}{c|ccccc}
Reference Foil                     & \ce{^{nat}In}         & \ce{^{nat}In}       & \ce{^{nat}In} & \ce{^{nat}In}     & \ce{^{nat}In}  \\ \hline
Reference Foil Mass (g)            & 0.248         & 0.248       & 0.241 & 0.247     & XXXXXX \\ \hline
Target Foil                        & \ce{^{nat}Zn}         & \ce{^{nat}Zn}       & \ce{^{nat}Zn} & \ce{^{nat}Ti}     & \ce{^{nat}Ti}  \\ \hline
Target Foil Mass (g)               & 0.538         & 0.451       & 0.452 & 0.337     & XXXX   \\ \hline
Irradiation Time, $t_1$ (s)           & 10800         & 10800       & 12629 & 11837     & XXXX   \\ \hline
Start of Count, $t_2$ (s)             & 12585         & 26985       & 14919 & 101245    & XXXX   \\ \hline
End of Count, $t_3$ (s)               & 103773        & 80993       & 68919 & 187669    & XXXX   \\ \hline
Photopeak Counts, 336 keV (\ce{^{115m}In}) & 113665 $\pm$ 1490 & 74321 $\pm$ 275 & XXXXX & 2122 $\pm$ 55 & XXXX   \\ \hline
Photopeak Counts, 391 keV (\ce{^{113m}In}) & 3382 $\pm$ 171    & 860 $\pm$ 40    & XXXXX &    \hrulefill       &  \hrulefill      \\ \hline
Photopeak Counts, 511 keV (\ce{^{64}Cu})   & 16055 $\pm$ 643   & 12852 $\pm$ 118 & XXXXX &    \hrulefill       &   \hrulefill     \\ \hline
Photopeak Counts, 159 keV (\ce{^{47}Sc})   &   \hrulefill            &    \hrulefill         &   \hrulefill    & 3877 $\pm$ 83 & XXXX  
\end{tabular}%
}
\end{table*}

% % Please add the following required packages to your document preamble:
% % \usepackage{booktabs}
% % \usepackage{graphicx}
% \begin{table*}
% \centering
% \caption{Counting times and photopeak counts for each of the (Zn/In) and (Ti/In)  experiments.}
% \label{tab:peak_counts}
% \resizebox{\textwidth}{!}{%
% \begin{tabular}{@{}|c|ccccc@{}}
% Reference Foil                     & natIn         & natIn       & natIn & natIn     & natIn  \\
% Reference Foil Mass (g)            & 0.248         & 0.248       & 0.241 & 0.247     & XXXXXX \\
% Target Foil                        & natZn         & natZn       & natZn & natTi     & natTi  \\
% Target Foil Mass (g)               & 0.538         & 0.451       & 0.452 & 0.337     & XXXX   \\
% Irradiation Time, t1 (s)           & 10800         & 10800       & 12629 & 11837     & XXXX   \\
% Start of Count, t2 (s)             & 12585         & 26985       & 14919 & 101245    & XXXX   \\
% End of Count, t3 (s)               & 103773        & 80993       & 68919 & 187669    & XXXX   \\
% Photopeak Counts, 336 keV (115mIn) & 113665 $\pm$ 1490 & 74321 $\pm$ 275 & XXXXX & 2122 $\pm$ 55 & XXXX   \\
% Photopeak Counts, 391 keV (113mIn) & 3382 $\pm$ 171    & 860 $\pm$ 40    & XXXXX &           &        \\
% Photopeak Counts, 511 keV (64Cu)   & 16055 $\pm$ 643   & 12852 $\pm$ 118 & XXXXX &           &        \\
% Photopeak Counts, 159 keV (47Sc)   &               &             &       & 3877 $\pm$ 83 & XXXX  
% \end{tabular}%
% }
% \end{table*}

\subsection{Experimental verification of incident neutron energy}\label{sec:lee_sec}

As shown in \autoref{sec:sample_loading} above, the effective neutron energy depends on the angle range subtended by the sample with respect to the incident deuteron beam.  In order to determine this angle it is necessary to measure the lateral location of the beam with respect to the sample location.  This centroid of the beam was measured using a 3 x 3 array of 0.5 cm diameter indium foils.  The relative activity of these foils was then determined via post-irradiation counting of the \ce{^{115m}In} isomer ($t_{1/2}$ = 4.486 h).  \autoref{fig:in_heat_plot} below shows the measured activities for these 9 indium foils.  Based on these values we are able to verify that the beam was indeed centered on the middle of the zinc and titanium samples.

% 
% \begin{figure}
%  \centering
%  \includegraphics[scale=0.6]{./figures/XXX.pdf}
%  % fracplot.pdf: 372x295 pixel, 72dpi, 13.12x10.41 cm, bb=0 0 372 295
%  \caption{\textcolor{red}{INSERT CAPTION HERE}}
%  \label{fig:in_heat_plot}
% \end{figure}
%

\comment{cire relevant NDS article for t1/2}

\comment{CHECK WITH ERIC for 3x3 In info and plot}

\subsection{Calculation of measured cross sections}\label{sec:calcs_sec}

For a thin target consisting of \(N_T\) target nuclei (with a reaction cross section $\sigma\pp{\bar{E}}$), subjected to a constant neutron flux $\phi\pp{\bar{E}}$, the rate of production ($R$) of the product nucleus will be:

\begin{equation}
R = N_T \sigma\pp{\bar{E}} \phi\pp{\bar{E}} 
\end{equation}

If this irradiation lasts for  some time $t_1$, and gamma ray spectrum acquisition begins at a later time $t_2$ and ends at  $t_3$, then the number of product decays ($D$; with decay constant $\lambda$) during the acquisition will be:

\begin{equation}
\begin{aligned}
D &= \dfrac{R}{\lambda}\pp{e^{\lambda t_1}-1}\pp{e^{-\lambda t_3} - e^{-\lambda t_2}}
\\
&= \dfrac{N_T \sigma\pp{\bar{E}} \phi\pp{\bar{E}} }{\lambda}\pp{e^{\lambda t_1}-1}\pp{e^{-\lambda t_3} - e^{-\lambda t_2}} 
\end{aligned}
\end{equation}

If this decay emits a gamma ray with branching ratio $I_\gamma$ (photons emitted per decay), and is detected with an absolute efficiency  of $\epsilon_\gamma$ (photons detected / photons emitted), then the number of observed decays between $t_2$ and $t_3$ will be:

% \begin{equation}
% \begin{aligned}
% N_{obs} &= D \epsilon_\gamma I_\gamma \\
% &= \? \epsilon_\gamma I_\gamma  \dfrac{N_T \sigma\pp{\bar{E}} \phi\pp{\bar{E}} }{\lambda}\pp{e^{\lambda t_1}-1}\pp{e^{-\lambda t_2} - e^{-\lambda t_3}} 
% \end{aligned}
% \end{equation}

\begin{align}
N_{obs} &= D \epsilon_\gamma I_\gamma \\
&=  \epsilon_\gamma I_\gamma  \dfrac{N_T \sigma\pp{\bar{E}} \phi\pp{\bar{E}} }{\lambda}\pp{e^{\lambda t_1}-1}\pp{e^{-\lambda t_3} - e^{-\lambda t_2}} \nonumber
\end{align}

Solving this equation for the  cross-section results in:

\begin{equation}\label{eqn:single_xs_eqn}
\sigma\pp{\bar{E}} = \dfrac{N_{obs}\lambda}{N_T \epsilon_\gamma I_\gamma  \phi\pp{\bar{E}}  \pp{e^{\lambda t_1}-1}\pp{e^{-\lambda t_3} - e^{-\lambda t_2}}}
\end{equation}

\autoref{eqn:single_xs_eqn} can be used to determine the unknown (n,p) cross sections relative to the well-known \ce{^{115}In}(n,n')\ce{^{115m}In} and \ce{^{113}In}(n,n')\ce{^{113m}In} inelastic scattering cross sections since the Zn and Ti samples were co-irradiated with indium foils. This approach has a number of advantages since the result is independent of neutron flux and only depends on the relative detector efficiencies at each gamma-ray energy.  \autoref{eqn:calc_eqn} shows the ratio of the cross sections determined using this approach, in which subscript $P$ indicates a quantity for either \ce{^{64}Cu} or \ce{^{47}Sc}, and subscript $In$ indicates a quantity for either the \ce{^{113m}In} or \ce{^{115m}In} isomer. A minor term was added to correct for the small self-attenuation of the gamma rays emitted by the activated foils:

\begin{align}\label{eqn:calc_eqn}
\dfrac{\sigma_P}{\sigma_{In}} &=  \dfrac{A_P}{A_{In}} \dfrac{N_{T,In}}{N_{T,P}} \dfrac{\lambda_P}{\lambda_{In}} \dfrac{e^{\lambda_{In}t_1}-1}{e^{\lambda_{P}t_1}-1}  \\
&\times \dfrac{e^{-\lambda_{In}t_3}-e^{-\lambda_{In}t_2}}{e^{-\lambda_{P}t_3} - e^{-\lambda_{P}t_2}} \dfrac{\epsilon_{In}}{\epsilon_P}  \dfrac{I_{\gamma,In}}{I_{\gamma,P}} \dfrac{e^{-\mu_{In}x_{In}/2}\times e^{-\mu_{In}x_{P}}}{e^{-\mu_{P}x_{P}/2}} \nonumber
\end{align}

where:

\begin{itemize}

\item $A$ is the integrated counts under a photopeak [counts],

\item $\sigma$ is the cross section for either the production of a product or isomer [mb],

\item $N_T$ is the initial number of target nuclei,

\item $\lambda$  is the decay constant [s$^{-1}$],

\item $t_1$ is the irradiation time [s],

\item $t_2$ is the time between the start-of-beam and the start of counting [s],

\item $t_3$ is the time  between the start-of-beam and the end of counting [s],

\item $\epsilon$ is the  detector efficiency for a particular photopeak,

\item $I_\gamma$ is the decay gamma ray branching ratio,

\item $\mu$ is the photon attenuation coefficient for a particular decay gamma ray in a foil [cm$^{-1}$],

\item and $x$ is the thickness of foil traversed by a particular decay gamma ray [cm] 
\end{itemize}

In addition to the \ce{^{115}In}(n,n')\ce{^{115m}In} reference cross section, the \ce{^{115}In}(n,$\gamma$)\ce{^{116m}In}  ($t_{1/2}$ = 54.29 min) activity can be used to determine the  \ce{^{64}Zn}(n,p) and \ce{^{47}Ti}(n,p) cross section.  The capture activity is potentially subject to contamination from lower energy, especially thermal, \enquote{room return} neutrons since the (n,$\gamma$) cross section at 25 meV is approximately YYY MeV greater than at 2.7 MeV.

\comment{cite relevant NDS article}

\comment{add appropriate numbers to (n,g) cross sections}

\comment{cite IRDFF somewhere?}

With the exception of decay constants, which have negligible uncertainty compared to other sources of uncertainties in this work, each of the parameters in this model carries a statistical uncertainty. Based on the assumption that these uncertainties are independent, the total relative statistical uncertainty $\delta_\sigma$ is calculated by taking the quadrature sum of the relative uncertainties of each parameter  $\delta_i$:

\begin{equation}
\delta_\sigma = \norm{ \ensuremath{\vec{\delta}} }_2 = \sqrt{\sum_{i=1}^N  \delta_i^2  }
\end{equation}

This total statistical uncertainty will be plotted as the cross-section error bar in the excitation function for these production reactions.

\subsection{Systematic errors}

The largest source of systematic uncertainty in the cross section determined via the \enquote{ratio approach} is the \textcolor{red}{YYY\%} uncertainty in the \ce{^{115}In}(n,n')\ce{^{115m}In}  cross section and the \textcolor{red}{YYY\%} uncertainty in the \ce{^{113}In}(n,n')\ce{^{113m}In}  cross section.  An additional uncertainty arises from the fact that the Zn/Ti samples are not located at exactly the same location as the indium monitor foils, and are therefore not subject to precisely the same neutron flux.  However, the MCNP6 simulations shown in \autoref{fig:mcnp_flux} indicate that the difference in the flux that the two foils are subjected to is less than \textcolor{red}{YY\%}.  Other monitor foils could be used instead of indium, with \ce{^{57}Ni}(n,p)\ce{^{57}Co} being one possible candidate, but the 4.486 hour and 99.476 minute half-lives of the \ce{^{115m}In} and \ce{^{113m}In} isomers \textcolor{red}{[XXXref NDS]}, respectively, makes indium a better candidate for measuring the production of radionuclides with lifetimes much less than 77+ days. The largest source of uncertainty in energy window arises from uncertainties in the actual dimension of the deuteron beam on the production target.  We believe, based on \enquote{burn marks} on the production target that the beam was approximately circular, with a flat intensity profile and a 5 mm diameter.  However, every 1 mm change in the beam radius would cause a \textcolor{red}{XXX} MeV shift in the centroid and a \textcolor{red}{YYY} MeV change in the effective energy bin.

\comment{filling in missing data}

\comment{Use your energy-angle figure from Lisken above to estimate these numbers. You can point to the figure in the text if you'd like.}

\section{Results}

Using the ratio method described here, the cross sections for the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} and \ce{^{64}Zn}(n,p)\ce{^{64}Cu} reactions have been calculated for incident neutron energy of 2.7645 MeV,  + 0.0151 MeV / - 0.0219 MeV. These values are recorded in \autoref{tab:xs_results}.

\comment{Need Mauricio's Cu point w/ more sigfigs. Need Mauricio's Sc point too}

\comment{Truncate sig figs here - Note that the systematic uncertainty you calculated in the previous paragraph will provide a natural limit on the number of significant figures you quote for the energy window.}

\comment{Report a weighted average between the solo values, insert in text}

% 
% % Please add the following required packages to your document preamble:
% % \usepackage{booktabs}
% \begin{table}[]
% \centering
% \caption{My caption}
% \label{my-label}
% \begin{tabular}{@{}cc@{}}
% \toprule
% h                                 & $\sigma$($E_n$ = 2.7645 MeV) (mb)                                                                     \\ \midrule
% \ce{^{64}Zn}(n,p)\ce{^{64}Cu} (relative to \ce{^{113}In}) & \begin{tabular}[c]{@{}c@{}}45.953 $\pm$ 3.351, \\ 46.493 $\pm$ 2.805, \\ 46.9 $\pm$ 3.189\end{tabular}  \\
% \ce{^{64}Zn}(n,p)\ce{^{64}Cu} (relative to \ce{^{115}In}) & \begin{tabular}[c]{@{}c@{}}49.716 $\pm$ 3.335, \\ 49.011 $\pm$ 2.698, \\ XXXX $\pm$ XXXXXX\end{tabular} \\
% \ce{^{47}Ti}(n,p)\ce{^{47}Sc} (relative to \ce{^{115}In}) & \begin{tabular}[c]{@{}c@{}}25.901 $\pm$ 1.7089, \\ XXXXX $\pm$ XXXXX\end{tabular}                   \\ \bottomrule
% \end{tabular}
% \end{table}

% Please add the following required packages to your document preamble:
% \usepackage{booktabs}
\begin{table}
\centering
\caption{Results of cross section measurement. Note that the last data point for the \ce{^{47}Sc} measurement (marked with *) was performed at a slightly different beam spot location, leading to a difference in effective neutron energy.}
\label{tab:xs_results}
\begin{tabular}{@{}cc@{}}
\toprule
Reaction                                                                            & $\sigma$($E_n$ = 2.7645 MeV) (mb)                                                                       \\ \midrule
\begin{tabular}[c]{@{}c@{}}\ce{^{64}Zn}(n,p)\ce{^{64}Cu} \\ (relative to \ce{^{113}In})\end{tabular} & \begin{tabular}[c]{@{}c@{}}45.953 $\pm$ 3.351, \\ 46.493 $\pm$ 2.805, \\ 46.9 $\pm$ 3.189\end{tabular}  \\
\begin{tabular}[c]{@{}c@{}}\ce{^{64}Zn}(n,p)\ce{^{64}Cu} \\ (relative to \ce{^{115}In})\end{tabular} & \begin{tabular}[c]{@{}c@{}}49.716 $\pm$ 3.335, \\ 49.011 $\pm$ 2.698, \\ XXXX $\pm$ XXXXXX\end{tabular} \\
\begin{tabular}[c]{@{}c@{}}\ce{^{47}Ti}(n,p)\ce{^{47}Sc} \\ (relative to \ce{^{115}In})\end{tabular} & \begin{tabular}[c]{@{}c@{}}25.901 $\pm$ 1.7089, \\ XXXXX $\pm$ XXXXX\end{tabular}                   \\ \bottomrule
\end{tabular}
\end{table}

\comment{Truncate sig figs in En in \autoref{tab:xs_results} }

The results for the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} and \ce{^{64}Zn}(n,p)\ce{^{64}Cu} production cross sections can also be seen plotted in Figures \ref{fig:sc_xs} and \ref{fig:cu_xs}, respectively. It is apparent that the relative cross sections measured in the present work are consistent with both the small quantity of other literature results, and are measured to within 6\% uncertainty, lower than existing measurements. In addition, the results presented in this work show strong agreement with the recommended values from both the ENDF/B-VII.1 nuclear data library and reaction modeling results from TALYS\cite{koning2005talys}.

\comment{Waiting on Lee for EMPIRE, ALICE data and language discussing differences and the significance of this.}

\comment{Mention Shimizu data specifically? }

\comment{Add a zoomed-in view like in \autoref{fig:cu_xs}?}

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/47Sc_fixed.pdf}
 % 47Sc_fixed.pdf: 375x294 pixel, 72dpi, 13.23x10.37 cm, bb=0 0 375 294
 \caption{Measured \ce{^{47}Ti}(n,p)\ce{^{47}Sc} relative production cross section.}
 \label{fig:sc_xs}
\end{figure}

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/64cu_fixed.pdf}
 % 47Sc_fixed.pdf: 375x294 pixel, 72dpi, 13.23x10.37 cm, bb=0 0 375 294
 \caption{Measured \ce{^{64}Zn}(n,p)\ce{^{64}Cu} relative production cross section}
 \label{fig:cu_xs}
\end{figure}

\begin{figure}
 \centering
 \includegraphics[scale=0.6]{./figures/64cu_fixed_zoom.pdf}
 % 47Sc_fixed.pdf: 375x294 pixel, 72dpi, 13.23x10.37 cm, bb=0 0 375 294
 \caption{A zoomed-in view of \autoref{fig:cu_xs}.}
 \label{fig:cu_xs_zoom}
\end{figure}

\section{Summary}

Using direct activation methods for thin foils as described in this work, the \ce{^{47}Ti}(n,p)\ce{^{47}Sc} and \ce{^{64}Zn}(n,p)\ce{^{64}Cu} production cross sections have been measured for  2.7645 MeV neutrons, to within 6\% uncertainty. These measurements are consistent with experimental data and theoretical models,  and have been measured to lower uncertainty than existing measurements. The use of DD neutron generators can be an efficient pathway for the measurement of these low-energy (n,p) reaction channels, as well as a relative method used to normalize measurements at higher neutron energies. While it reduces the neutron flux incident upon target foils, the use of collimated beam lines to irradiate foils at various angles allows a single neutron generator to perform these measurements at various other energies throughout the DD neutron emission energy spectrum.

\comment{Future work? Thermal neutron measurements? Something like the following?}

 Future work will involve the continued measurement of the (n,p) production cross sections for various other emerging therapeutic and diagnostic radioisotopes, to expand the toolset of options available for modern medical imaging and cancer therapy. This will focus on radionuclides which permit more customized and precise dose deposition, as well as patient-specific treatments

 \comment{Should we specifically mention other isotopes we plan to measure?}

 \section{Acknowledgements}

 This work has been carried out at the University of California, Berkeley, and performed under the auspices of the U.S. Department of Energy by Lawrence Livermore National Laboratory under contract \# DE-AC52-07NA27344 and Lawrence Berkeley National Laboratory under contract \# DE-AC02-05CH11231. Funding has been provided from the US Nuclear Regulatory Commission and the US Nuclear Data Program.

 \comment{Who else should we acknowledge?}

% \pagebreak
% 
% \onecolumn
% 
% \appendix

% \section{Data} \label{data}
% 
% % \begin{longtable}{|c|c|c|c|c|} 
% % \caption{My caption}
% % \label{tab:dummy}
% % % \begin{tabular}{|c|c|c|c|c|}
% %    \hline
% %  
% % \end{longtable}
% hello world

% 
% \section{Results} \label{equations}
% 
% Tabulated results are listed here, to avoid breaking text flow in the main text.
% 
% \begin{table}[h]
% \centering
% \caption{Transition conversion coefficients and efficiencies}
% \label{my-label2}
% \begin{tabular}{c|c|c|c|c}
% Nucleus & Transition       & E\(_\gamma\) (keV) & \(\alpha\), Conversion Coefficient  & \(\epsilon\), Intrinsic Efficiency \\ \hline
% \ce{^{110}Ru}   & 2\(\rightarrow\)0 & 240.7          & 0.0569           \(\pm\) 0.0008              & 1863.906   \\
%         & 4\(\rightarrow\)2 & 422.6          & 0.00887          \(\pm\) 0.00013             & 1219.507   \\
%         & 6\(\rightarrow\)4 & 575.7          & 0.00356          \(\pm\) 0.00005             & 956.569    \\ \hline
% \ce{^{140}Xe}   & 2\(\rightarrow\)0 & 376.657        & 0.0205          \(\pm\) 0.0003              & 1340.054   \\
%         & 4\(\rightarrow\)2 & 457.63         & 0.01154          \(\pm\) 0.00017             & 1142.505   \\
%         & 6\(\rightarrow\)4 & 582.4          & 0.00593          \(\pm\) 0.00009             & 948.705    \\ \hline
% \ce{^{140}Ba}   & 2\(\rightarrow\)0 & 602.35         & 0.00599         \(\pm\) 0.00009             & 926.624    \\
%         & 4\(\rightarrow\)2 & 528.25         & 0.00848         \(\pm\) 0.00012             & 1019.604   \\
%         & 6\(\rightarrow\)4 & 529.7          & 0.00842         \(\pm\) 0.00012             & 1017.470   \\ \hline
% \ce{^{144}Ba}   & 2\(\rightarrow\)0 & 199.326        & 0.175           \(\pm\) 0.0025              & 2085.452   \\
%         & 4\(\rightarrow\)2 & 330.88         & 0.0333         \(\pm\) 0.0005              & 1485.782   \\
%         & 6\(\rightarrow\)4 & 431.3          & 0.015          \(\pm\) 0.00021             & 1199.258  
% \end{tabular}
% \end{table}
% 
% \begin{table}[h]
% \centering
% \caption{Extracted transition intensities}
% \label{my-label}
% \begin{tabular}{c|c|c|c|c|c}
% Nucleus & Transition       & E\(_\gamma\) (keV) & \(A_{obs}\), Peak Area (a.u.)   & \(A_{source}\),  Peak Intensity (a.u.)  & \(R_i\), Relative Intensity  \\ \hline
% \ce{^{110}Ru}   & 2\(\rightarrow\)0 & 240.7          & 1097    \(\pm\)   136        & 0.622          \(\pm\) 0.077             & 1.000              \(\pm\) 0.175                 \\
%         & 4\(\rightarrow\)2 & 422.6          & 515     \(\pm\)   107        & 0.426          \(\pm\) 0.089             & 0.685              \(\pm\) 0.200                 \\
%         & 6\(\rightarrow\)4 & 575.7          & 144      \(\pm\)  55         & 0.151          \(\pm\) 0.058             & 0.243              \(\pm\) 0.198                 \\ \hline
% \ce{^{140}Xe}   & 2\(\rightarrow\)0 & 376.657        & 542     \(\pm\)   129        & 0.413          \(\pm\) 0.098             & 1.000              \(\pm\) 0.337                 \\
%         & 4\(\rightarrow\)2 & 457.63         & 357       \(\pm\) 59         & 0.316          \(\pm\) 0.052             & 0.766              \(\pm\) 0.254                 \\
%         & 6\(\rightarrow\)4 & 582.4          & 161       \(\pm\) 61         & 0.171          \(\pm\) 0.065             & 0.414              \(\pm\) 0.288                 \\ \hline
% \ce{^{140}Ba}   & 2\(\rightarrow\)0 & 602.35         & 822       \(\pm\) 164        & 0.892          \(\pm\) 0.178             & 1.000              \(\pm\) 0.282                 \\
%         & 4\(\rightarrow\)2 & 528.25         & 384       \(\pm\) 78         & 0.380          \(\pm\) 0.077             & 0.426              \(\pm\) 0.186                 \\
%         & 6\(\rightarrow\)4 & 529.7          & 310       \(\pm\) 104        & 0.307          \(\pm\) 0.103             & 0.344              \(\pm\) 0.229                 \\ \hline
% \ce{^{144}Ba}   & 2\(\rightarrow\)0 & 199.326        & 900       \(\pm\) 242        & 0.507          \(\pm\) 0.136             & 1.000              \(\pm\) 0.380                 \\
%         & 4\(\rightarrow\)2 & 330.88         & 535       \(\pm\) 127        & 0.372          \(\pm\) 0.088             & 0.734              \(\pm\) 0.307                 \\
%         & 6\(\rightarrow\)4 & 431.3          & 363       \(\pm\) 64         & 0.307          \(\pm\) 0.054             & 0.606              \(\pm\) 0.250                
% \end{tabular}
% \end{table}
%     
% \pagebreak
% 
% 
% 
% \section{Supplemental Figures} \label{fit_figures}
% 
% The remaining figures of the fitted \(\gamma\)-ray transition photopeaks are included here for reference, to avoid over-cluttering the main text.
% 
% 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit110Ru240-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{110}Ru} 240.7 keV peak and its surroundings.}
% %  \label{fig:110Ru240}
% % \end{figure}
% % 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit110Ru422-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{110}Ru} 422.6 keV peak and its surroundings.}
% %  \label{fig:110Ru422}
% % \end{figure}
% % 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit110Ru575-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{110}Ru} 575.7 keV peak and its surroundings.}
% %  \label{fig:110Ru575}
% % \end{figure}
% 
% 
% 
% 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit140Xe376-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{140}Xe} 376.7 keV peak and its surroundings.}
% %  \label{fig:140Xe376}
% % \end{figure}
% % 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit140Xe457-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{140}Xe} 457.6 keV peak and its surroundings.}
% %  \label{fig:140Xe457}
% % \end{figure}
% % 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit140Xe582-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{140}Xe} 582.3 keV peak and its surroundings.}
% %  \label{fig:140Xe582}
% % \end{figure}
% % 
% % 
% % 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit140Ba602-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{140}Ba} 602.4 keV peak and its surroundings.}
% %  \label{fig:140Ba602}
% % \end{figure}
% % 
% % 
% % 
% % 
% % 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit144Ba199-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{144}Ba} 199.3 keV peak and its surroundings.}
% %  \label{fig:144Ba199}
% % \end{figure}
% % 
% % 
% % 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit144Ba330-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{144}Ba} 330.9 keV peak and its surroundings.}
% %  \label{fig:144Ba330}
% % \end{figure}
% % 
% % \begin{figure}[h]
% %  \centering
% %  \includegraphics[scale=0.7]{./hw04/fit144Ba431-cropped.pdf}
% %  % fit110Ru240.ps: 570x750 pixel, 72dpi, 20.11x26.46 cm, bb=0 0 570 750
% %  \caption{Fit to the \ce{^{144}Ba} 431.3 keV peak and its surroundings.}
% %  \label{fig:144Ba431}
% % \end{figure}
% 
% 
% 
%     
% \pagebreak
% 
% 
% % \appendix
% \section{Source Code} \label{source}
% 
% \begin{center}
%  \texttt{gint.py} - Implementation of the Wilhelmy statistical model, used to estimate total transition intensities in the ground state band  of primary fission fragments
% \end{center}
% 
% % \begin{lstlisting}
% % \lstloadlanguages{Python}
% 
% % \lstinputlisting[linewidth=\columnwidth,language=Python,breaklines=true,frame=single]{./hw04/gint.py}
% % % \end{lstlisting}
% 
% 
% 
% \begin{center}
%  \texttt{hw04\_04.m} - used to process and plot statistical transition intensity curves
% \end{center}
% 
%  
% %   \lstinputlisting[frame=single]{./hw04/hw04_04.m}
% 
% 
% % \begin{verbatim}[commandchars=\\\{\}]
% % \verbatiminput{./HW01/test.txt}
% % \end{verbatim}
% 
% %% The Appendices part is started with the command \appendix;
% %% appendix sections are then done as normal sections
% %% \appendix
% 
% %% \section{}
% %% \label{}
% 
% %% References
% %%
% %% Following citation commands can be used in the body text:
% %% Usage of \cite is as follows:
% %%   \cite{key}         ==>>  [#]
% %%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

% \twocolumn

%% References with BibTeX database:

% \bibliographystyle{elsarticle-num}
% \bibliographystyle{elsarticle-harv}
% \bibliographystyle{elsarticle-num-names}
% \bibliography{<your-bib-database>}
% \addcontentsline{toc}{chapter}{Bibliography}
% \bibliographystyle{elsarticle-num}
\bibliographystyle{ieeetr}
\bibliography{../../library}
% \thispagestyle{fancyTOC}

H. F. Aly et al., Microchim. Acta, vol. 59, no. 1, 1971.

K. S. Bhatki et al., J. Radioanal. Chem., vol. 2, no. 1-2, 1969.

T. H. Bokhari et al., J. Radioanal. Nucl. Chem., vol. 283, no. 2, 2010.

J. F. Briesmeister et al., Los Alamos National Laboratory, 1986.

M. B. Chadwick,et al., Nucl. Data Sheets, vol. 107, no. 12, 2006.

A. J. Koning et al., AIP Conference Proceedings, 2005, vol. 769, no. 2.

H. Liskien et al., Nucl. Data Tables, vol 11, 1973.

M. R. Lewis et al., J. Nucl. Med., vol. 44, no. 8, Aug. 2003.

C. Müller et al., J. Nucl. Med., vol. 55, no. 10, Oct. 2014.

N. Otuka et al., Nucl. Data Sheets, vol. 120, 2014.

L. Pietrelli et al., J. Radioanal. Nucl. Chem., vol. 157, no. 2, 1992.

S. M. Qaim et al., IAEA Technical Reports Series No. 473, 2011.

T. Shimizu, et al, Ann. Nucl. Energy, vol. 31, no. 9, pp. 975-990, 2004.

T. Shimizu, et al, Nucl. Instruments Methods Phys. Res. Sect. A Accel. Spectrometers, Detect. Assoc. Equip., vol. 527, no. 3, pp. 543-553, 2004.

D. Updegraff et al., ''Nuclear Medicine without Nuclear Reactors or Uranium Enrichment,'' 2013.

\comment{Cite Cory's thesis, or upcoming NIM paper?}

\comment{Convert these to BibTeX format!}

%% Authors are advised to use a BibTeX database file for their reference list.
%% The provided style file elsarticle-num.bst formats references in the required Procedia style

%% For references without a BibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\end{document}

%%
%% End of file `ecrc-template.tex'.